---
title: "Reproducible Research"
author: "Ruíz Rodríguez Ricardo"
date: "26/9/2020"
output:
  html_document: default
  pdf_document: default
---

## Loading and preprocessing the data

```{r simulation, echo = TRUE}
  activity <- read.csv("activity.csv")
```


```{r,  echo = TRUE}
  str(activity)
```

```{r , echo = TRUE}
  summary(activity)
```

## What is mean total number of steps taken per day?

First load the packages dplyr and ggplot2

```{r, results='hide'}
  library(dplyr)
  library(ggplot2)
```

1.- Calculate the total number of steps taken per day

```{r , echo=TRUE}
  activity_day <- activity%>%
  group_by(date)%>%
  summarise(steps = sum(steps, na.rm = T))
  head(activity_day)
```

2.- Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
  with(activity_day, ggplot(mapping = aes(steps))+
       geom_histogram(col='black', fill='green', alpha = 0.4))

```


3.-Calculate and report the mean and median of the total number of steps taken per day

Here the variable act_mm_d and act_mmed_d, its represent the  average activity  by day

```{r, echo=TRUE}
  act_mm_d <- mean(tapply(activity$steps, activity$date, sum), na.rm = T)

  act_mmed_d <- median(tapply(activity$steps, activity$date, sum), na.rm = T)

```


## What is the average daily activity pattern?

1.- Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days

```{r, echo=TRUE}
  
  activity.avgstep <- tapply(activity$steps, activity$interval, mean, na.rm = T)

  plot(activity.avgstep, type = "l", main = "The average steps for intervals of 5 minuts",
     xlab = "interval of 5 minutes")
```


2.-Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

  First number that we look at here down, is the maximun number of steps for the month octouber and the 
  second number is the maximun number of steps for the month November

```{r, echo=TRUE}
  maxstep <- which.max(activity.avgstep)
  maxstep
```

## Imputing missing values

1.- Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs.

  I calculated the total number of de registers NAs with the  summary function, we notated that the colum steps is only that contained this values Na.
  
```{r, echo=TRUE}
  summary(activity)
```

2.- The next part i devise strategy where  i'll switch the values Na by the mean  the 5-minute interval steps.

I do used library Hmisc for switching  the values NA of the colum STEP

```{r, echo=FALSE}
  library(Hmisc)
```

I created a copy of activity and change the NA values by the average of the total data. 

```{r, echo=TRUE}
  activity.copy <- activity

  activity.copy$steps <- impute(activity.copy$steps, mean)

  summary(activity.copy)
```


3.- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.


```{r, echo=TRUE}
  activity.copy.total.steps <- activity.copy%>%
  group_by(date)%>%
  summarise(total.steps = sum(steps))

  with(activity.copy.total.steps,ggplot(mapping = aes(total.steps))+
    geom_histogram(col='black', fill='green', alpha = 0.4))

```


## Are there differences in activity patterns between weekdays and weekends?

1.- Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

I'll use the pakage lubridate to compared the average number steps taken per 5-minute interval across weekdays and weekends

```{r}
  library(lubridate)

  activity.copy$date <- as.POSIXct(activity.copy$date)

```


In this part, i added a colum to the copy data frame, called week, this variable contain the week numbers, for example: wednesday is te number 4, sunday is the number 1, etc  

```{r, echo=TRUE}
  activity.copy <- mutate(activity,week = wday(activity$date))

  activity.weekdays<-activity.copy%>%
    filter(week == 2 | week == 3 | week == 4 | week == 5 | week == 6)

  activity.weekends<-activity.copy%>%
    filter(week == 7 | week == 1)
```

It calculate mean of the two colums, by the  weekdays and by the weekends.

```{r, echo=TRUE}
  act.weekday.mean <- tapply(activity.weekdays$steps, activity.weekdays$interval, mean, na.rm = T)

 act.weekend.mean <- tapply(activity.weekends$steps, activity.weekends$interval, mean, na.rm = T)
```

Created the comparation inter the two histograms

```{r}
  par(mfrow = c(2,1), mar = c(4,4,2,1))

  plot(act.weekday.mean, type = "l", xlab = "inteval", ylab = "average steps",
     main = "activity of each weekday")

  plot(act.weekend.mean, type = "l", xlab = "inteval", ylab = "average steps",
     main = "activity of each weekend")
```
